A mixture of hidden Markov models to predict the lymphatic spread in head and neck cancer depending on primary tumor location

### 2024-10-10 _tex/manuscript.tex
Built site for gh-pages

The mixture model assumes that the data is generated by a set of \(M\)
different underlying models. In our case, we introduce \(M\) distinct
lymphatic spread models. Each patient \(k\) is assumed to be generated
by one specific model from this set of \(M\) models.

@@

The latent variables \(\boldsymbol{\epsilon}_k\) are unknown and must be
inferred from the observed data. The mixture model assumes that each
patient is generated by exactly one of the \(M\) models. However, in our
application, this assumption does not reflect continuous or overlapping
behaviors between subsites. Instead, we infer a probabilistic
distribution over the latent variables, given the observed data and
model parameters. The posterior distribution:

@@

>>>>

The mixture model assumes that the data is generated by a set of \(M\)
different lymphatic spread models. Each patient \(k\) is then assumed to
be generated by one specific model from this set of \(M\) models.

@@

The latent variables \(\boldsymbol{\epsilon}_k\) are unknown and must be
inferred from the observed data. The mixture model assumes that each
patient is generated by exactly one of the \(M\) models. To do so a
probabilistic distribution over the latent variables, given the observed
data and model parameters is computed, the posterior distribution:

@@

It is important to mention, that the latent variables
\(\boldsymbol{\epsilon}\) are only used for parameter estimation. The
subsite specific mixture parameters \(\boldsymbol{\pi}^s\) are the only
parameters which are used for later risk predictions. In
figure~\ref{fig-model-simple} the mixture parameters are illustrated.
Subsites with different spread patterns, such as Gum (C03) and Base of
tongue (C01), are expected to get different model assignments.
Nonetheless, the latent variables for two patients with the same
diagnosis, i.e.~same involvement \(\boldsymbol{\xi}\) and and t-stage
\(T\), but different subsites are the same.

\includegraphics{figures/mixture_model_simplified.png}

\caption{\label{fig-model-simple}Illustration of mixture parameter
assignment. Since Gum and Base of tongue express different spread
patterns, the two models are expected to have different model
assignments. The arrow visibility represents the value of the mixture
parameter \(\pi\), where the more visible the arrow, the larger the
value for \(\pi\)}
